---
title: "Project Plan"
author: "Kingsley David Onyewuchi Ofoegbu, Jazelle Pilao, Doudou Shi, and Evan Steinberg"
date: "Due 2020-10-07" 
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Title of Project and Group members
## Title
Exploring how Coronavirus has Affected Individuals by State

## Group Members
Kingsley David Onyewuchi Ofoegbu, Jazelle Pilao, Doudou Shi, and Evan Steinberg.

# Description of the problem to be analyzed in the Shiny App

In this project, we are creating a Shiny app that explores coronavirus in the United States. Specifically, we are interested in creating an app that allows users to explore causes of coronavirus cases in the United States. Furthermore, we are interested in creating an app that can determine if there are any difference between identity groups in the United States when it comes to coronavirus. To explore this idea, we are looking at comparing  states, age groups, and pre-existing medical conditions. In addition, we are also interested in giving the user the option of comparing those groups by the total number of cases, number of positive/negative test results, number of active cases, and total deaths. Therefore, the user will have the ability to select the group that they are interested in exploring, as well as what will be compared. These comparisons will be done both visually and via statistical analysis.

# Proposed Data Sources

Our first proposed source is from The COVID Tracking Project found [here](https://covidtracking.com/data/download). This source provides information about the pandemic from January until the day before downloading. This ensures that the date is up to date at all times. This source gives data on confirmed deaths, confirmed hospitalizations, number of people on ventilators in that state, number of positive/negative COVID tests, etc. 

Our second proposed source is from HealthData.gov and can be found [here](https://healthdata.gov/dataset/conditions-contributing-deaths-involving-coronavirus-disease-2019-covid-19-age-group-and). This contains data from February to October. This source gives us information on coronavirus broken down by state. However, this also gives us information by age group, state, and any pre-existing conditions.

# Concept for Shiny App
## Overall layout

We understand that the overall layout might change after we learn about Shiny layouts in class. However, at this moment, we are interested in creating different tabs to house the different elements. One tab that we will create is an "about" tab, which explains the purpose of creating this app. In addition, we are interested in creating a visualization/exploration tab. In this tab, we will show all the various graphs/proportions that we will create (further described below). Finally, we are also interested in creating a statistical analysis tab where we show the analysis tests that we will be creating (more information on the exact analysis done below).

## Data Analysis and Visualization with User Choices and Results/Statistical Modeling with User Choices and Results

Individuals will have the opportunity to investigate coronavirus through exploring COVID positive results, progression to extreme medical attention (operationalized through individuals on ventilators), and death rates through several independent variables (age, location by state, pre-existing conditions). Individuals will have the opportunity to explore general population statistics by age, state, and pre-existing conditions in the United States before conducting further investigation. To allow individuals to investigate the effects of coronavirus on individual states, age groups, and individuals with different pre-existing conditions the following data visualization and statistical analysis will be available for individuals to conduct. First, people will have the opportunity to investigate the general proportion of individuals who had a confirmed COVID case to the general population and COVID-related deaths to the general population since February of this year. Next individuals will be able to visualize confirmed cases of COIVD by age range, state, and overall in the United States. Additionally, individuals will be able to look at the COVID-related death rates by age, state, effects of pre-existing conditions, and overall in the United States. All of these will be visualized through bar graphs and pie charts accordingly. After the option to visualize these proportional relations to the general population. Individuals will have the opportunity to visually investigate the relationship of age, existing conditions, and location (by state) to total COVID positive test results, individuals on ventilators, and COVID deaths. Individuals will have the opportunity to visualize these relationships through boxplots and a summary output will be generated for each visualization. Additional to the visualizations described above, individuals will have the opportunity to conduct several statistical analyses on the data previously described. Specifically, individuals will be able to conduct an ANOVA test to investigate if there is a statistically significant difference in COVID outcomes (user option of three groups: COVID positive cases, progression to extreme medical attention, and death rates) by age, location, and pre-existing condition diagnosis. Visualization for this analysis will also be created through box-plots. Summary of visualizations and statistical analysis by user input and results can be found below.

``` {r, echo = FALSE}
knitr::kable(tibble::tribble(
  ~"Topic of Investigation", ~"User Input Options", ~"Result Visualization",
  "General Unaffected Proportions", "Age, Location by State, Pre-existing Conditions", 
  "Pie Chart", "COVID Proportions", "Positive cases, progression to extreme medical intervention, total deaths", "Pie Chart/Bar Chart", "COVID outcomes by group", 
  "GROUPS: Age, Location by State, Pre-existing condition. OUTCOME: Positive cases, progression to extreme medical intervention, total deaths", 
  "Box-plots, Written Summary Statistics", "Statistical Difference in Means Analysis (ANOVA)",
  "GROUPED BY: Age, Location by State, Pre-existing Conditions. OUTCOMES MEASURED: Positive cases, progression to extreme medical intervention, total deaths", 
"Box-plot, Written Summary Statistics"

)
             , "simple")
```

# Steps/Schedule/Allocation of Responsabilities

``` {r, echo = FALSE}
knitr::kable(tibble::tribble(
  ~"Project Steps", ~"Complete By Date", ~"   Responsible Individuals",
  "Downlading/Cleaning Data", "11/01/2020", "   Evan, Doudou",
  "Creating Shared Repo", "11/01/2020", "   Kingsley",
  "About Page", "11/20/2020", "   Jazelle",
  "COVID Outcomes by Group Visualizations", "11/20/2020", "   Evan",
  "Proportion Visualizations", "11/20/2020", "   Doudou",
  "ANOVA Analysis and Visualization", "11/20/2020", "   Jazelle",
  "Aesthetics", "12/01/2020", "   Kingsley",
  "Writing up Vignette", "12/01/2020", "   Evan, Doudou, Jazelle, & Kingsley",
  "App Testing/Finalizing", "12/08/2020", "   Evan, Doudou, Jazelle, & Kingsley"
),
"simple")
```

# References

Center for Disease Control and Prevention. (2020, October 14). Conditions contributing to deaths involving coronavirus disease 2019 (COVID-19), by age group and state, United States. https://healthdata.gov/dataset/conditions-contributing-deaths-involving-coronavirus-disease-2019-covid-19-age-group-and. 

Our Data. The COVID Tracking Project. (2020). https://covidtracking.com/data/download. 
